### 2015-11-09 HSF-TN/draft-2015-LIC/draft-HSF-TN-2015-LIC.tex

We begin by defining some of the key terms as described in \cite{[3]}.
\begin{itemize}
\item Copyright is a legal right created by the law of a country that grants the creator of original work exclusive rights to its use and distribution, usually for a limited time. Copyright is a form of intellectual property, applicable to any expressed representation of a creative work. It is often shared among multiple authors, each of whom holds a set of rights to use or licence the work. These rights frequently include reproduction, control over derivative works, distribution, and "moral rights" such as attribution.

\item Public domain software is software that has been placed in the public domain. In other words there is absolutely no ownership such as copyright, trademark, or patent. Unlike other classes of licences, there are no restrictions as to what can be done with the software. The software can be modified, distributed, or sold even without any attribution. This is the simplest way to make open source software and allows people to share the program and their improvements, if they are so minded, but it also allows the program to be converted into proprietary software. They can make changes, many or few, and distribute the result as a proprietary product. People who receive the program in that modified form do not have the freedom that the original author gave them. 
A software licence is a legal instrument governing the use or redistribution of software. A typical software licence grants an end-user permission to use one or more copies of software in ways where such a use would otherwise potentially constitute copyright infringement of the software owner's exclusive rights under copyright law.

\item An open source software licence is a notice that grants the recipient of a piece of software extensive rights to modify and redistribute that software. Copyright law usually prohibits these actions, but the rights-holder (usually the author) of a piece of software can remove these restrictions by accompanying the software with a software licence that grants the recipient these rights. Software using such a licence can meet the conditions to be classed as open source software as conferred by the copyright holder. Open source licenses broadly divide into free software licenses and permissive software licenses.
Free software licences include "copyleft" provisions, which require all future versions to also be distributed with these freedoms. These are termed "restrictive" licences.

\item Permissive software licences do not impose these additional conditions and are usually just a grant of rights and a disclaimer of warranty, thus also allowing distributors to add restrictions for further recipients, or to produce an extended proprietary version of the software
All major open source software licences require that acknowledgement is given to authors of the software in documentation and/or at runtime. In an academic context these provisions can be useful in establishing the impact of a software project, and even when software released under a permissive licence is reused in a closed-source commercial product.

@@

The Copyleft principles were laid down by Richard Stallman of the Free Software Foundation in 1985, which was at the inception of the OSS movement through the creation of the GNU project. Copyleft is a general concept, and therefore cannot be used directly; you can only use a specific implementation of the concept. In the GNU Project, for example, the specific distribution terms used for most software are contained in the GNU General Public Licence (GPL) \cite{[4]}. GPL version 3 was published in 2007 but many copyleft projects (eg Linux) have chosen to continue with GPL version 2.
GPL 2 and 3 software amount to 24% and 10% of open source software respectively [6].

@@

As with full copyleft licences which prevent modified versions from being distributed under a proprietary licence, weak copyleft licenses are intended to ensure the non-appropriation by third parties of the Open Source software. 
As of July 2013, the LGPL was used by 7% of all open source licenced projects \cite{[6]}.

@@

BSD licences from 1988 onwards are another family of permissive free software licences, imposing minimal restrictions on the redistribution of covered software. Two variants of the licence, the New BSD Licence/Modified BSD Licence (3-clause), and the Simplified BSD Licence/FreeBSD Licence (2-clause) have been verified as GPL-compatible free software licences by the Free Software Foundation, and have been vetted as open source licences by the Open Source Initiative.
As of July 2013, the ASL, BSD and MIT permissive licences accounted for 42% of all open source licensed projects [6].

@@

The licence should contain a statement in the header of each source file acknowledging (i) the copyright of the owner(s) and (ii) the applicable licence.
\newline\verb?(i)	Copyright?

In the following we give some examples of Copyright statements that are used by CERN depending on whether the software is owned solely by CERN or by CERN and external partners:
\begin{itemize}
\item “© Copyright [year] CERN”
for software owned solely by a single institute, in this case CERN
\item “© Copyright [year] CERN [for the benefit of the [Name of appropriate group] Collaboration]”
for software developed by a collaboration but where ownership has been transferred to a single institute, in this case CERN
\item  “© Copyright [year] [names of all copyright holders]”
for software owned by partners in small collaborations
\item “© Copyright [year] Copyright Holders of [name of the collaboration or joint project]. See [https://link] for details of the Copyright Holders”
for software owned by partners in large collaborations
\end{itemize}

(ii)	Applicable licence

One of the following licence statements must be included (immediately following the copyright statement):

“This software is distributed under the terms of the GNU General Public Licence version 3 (GPL Version 3), copied verbatim in the file “COPYING” /copied verbatim below”.
For software distributed under the default GPLv3 licence
“This software is distributed under the terms of the GNU Lesser General Public Licence version 3 (LGPL Version 3), copied verbatim in the file “COPYING” /copied verbatim below”.
For software distributed under the LGPLv3 licence
“This software is distributed under the terms of the Apache version 2.0 licence, copied verbatim in the file “COPYING” /copied verbatim below”.
For software distributed under the Apache licence v2
“This software is distributed under the terms of the BSD-2-Clause licence, copied verbatim in the file “COPYING” /copied verbatim below”.
For software distributed under the BSD-2-Clause licence
“This software is distributed under the terms of the BSD-3-Clause licence, copied verbatim in the file “COPYING” /copied verbatim below”.
For software distributed under the BSD-3-Clause licence
“This software is distributed under the terms of the MIT licence, copied verbatim in the file “COPYING” /copied verbatim below”.
For software distributed under the MIT licence

>>>>

We begin by defining some of the key terms as described in \cite{[3]}.
\begin{itemize}
\item \textit{Copyright} is a legal right created by the law of a country that grants the creator of original work exclusive rights to its use and distribution, usually for a limited time. Copyright is a form of intellectual property, applicable to any expressed representation of a creative work. It is often shared among multiple authors, each of whom holds a set of rights to use or licence the work. These rights frequently include reproduction, control over derivative works, distribution, and "moral rights" such as attribution.

\item \textit{Public domain software} is software that has been placed in the public domain. In other words there is absolutely no ownership such as copyright, trademark, or patent. Unlike other classes of licences, there are no restrictions as to what can be done with the software. The software can be modified, distributed, or sold even without any attribution. This is the simplest way to make open source software and allows people to share the program and their improvements, if they are so minded, but it also allows the program to be converted into proprietary software. They can make changes, many or few, and distribute the result as a proprietary product. People who receive the program in that modified form do not have the freedom that the original author gave them. 
A software licence is a legal instrument governing the use or redistribution of software. A typical software licence grants an end-user permission to use one or more copies of software in ways where such a use would otherwise potentially constitute copyright infringement of the software owner's exclusive rights under copyright law.

\item An \textit{open source software licence} is a notice that grants the recipient of a piece of software extensive rights to modify and redistribute that software. Copyright law usually prohibits these actions, but the rights-holder (usually the author) of a piece of software can remove these restrictions by accompanying the software with a software licence that grants the recipient these rights. Software using such a licence can meet the conditions to be classed as open source software as conferred by the copyright holder. Open source licenses broadly divide into free software licenses and permissive software licenses.
Free software licences include "copyleft" provisions, which require all future versions to also be distributed with these freedoms. These are termed "restrictive" licences.

\item \textit{Permissive software licences} do not impose these additional conditions and are usually just a grant of rights and a disclaimer of warranty, thus also allowing distributors to add restrictions for further recipients, or to produce an extended proprietary version of the software
All major open source software licences require that acknowledgement is given to authors of the software in documentation and/or at runtime. In an academic context these provisions can be useful in establishing the impact of a software project, and even when software released under a permissive licence is reused in a closed-source commercial product.

@@

The Copyleft principles were laid down by Richard Stallman of the Free Software Foundation in 1985, which was at the inception of the OSS movement through the creation of the GNU project. Copyleft is a general concept, and therefore cannot be used directly; you can only use a specific implementation of the concept. In the GNU Project, for example, the specific distribution terms used for most software are contained in the GNU General Public Licence (GPL) \cite{[4]}. GPL version 3 was published in 2007 but many copyleft projects (eg Linux) have chosen to continue with GPL version 2.
GPL 2 and 3 software amount to 24\% and 10\% of open source software respectively [6].

@@

As with full copyleft licences which prevent modified versions from being distributed under a proprietary licence, weak copyleft licenses are intended to ensure the non-appropriation by third parties of the Open Source software. 
As of July 2013, the LGPL was used by 7\% of all open source licenced projects \cite{[6]}.

@@

BSD licences from 1988 onwards are another family of permissive free software licences, imposing minimal restrictions on the redistribution of covered software. Two variants of the licence, the New BSD Licence/Modified BSD Licence (3-clause), and the Simplified BSD Licence/FreeBSD Licence (2-clause) have been verified as GPL-compatible free software licences by the Free Software Foundation, and have been vetted as open source licences by the Open Source Initiative.

As of July 2013, the ASL, BSD and MIT permissive licences accounted for 42\% of all open source licensed projects [6].

@@

The licence should contain a statement in the header of each source file acknowledging the copyright of the owner(s) and the applicable licence.
\newline\verb?(i)	Copyright?

\subsection{Copyright}
\label{sec:CopyrightExamples}

In the following we give some examples of Copyright statements that are used by CERN depending on whether the software is owned solely by CERN or by CERN and external partners:
\begin{itemize}
\item for software owned solely by a single institute, in this case CERN:
\begin{quote}
© Copyright [year] CERN
\end{quote}
\item for software developed by a collaboration but where ownership has been transferred to a single institute, in this case CERN:
\begin{quote}
© Copyright [year] CERN [for the benefit of the [Name of appropriate group] Collaboration]
\end{quote}
\item for software owned by partners in small collaborations:
\begin{quote}
© Copyright [year] [names of all copyright holders]
\end{quote}
\item for software owned by partners in large collaborations:
\begin{quote}
© Copyright [year] Copyright Holders of [name of the collaboration or joint project]. See [https://link] for details of the Copyright Holders
\end{quote}
\end{itemize}

\subsection{Applicable licence}
\label{sec:ApplicableLicenceExamples}

One of the following licence statements must be included, immediately following the copyright statement, \textbf{and followed by the text of the relevant license as shown in the references}:

\begin{itemize}
\item For software distributed under the default GPLv3 licence\cite{[9]}:
\begin{quote}
This software is distributed under the terms of the GNU General Public Licence version 3 (GPL Version 3).
\end{quote}
\item For software distributed under the LGPLv3 licence\cite{[10]}:
\begin{quote}
This software is distributed under the terms of the GNU Lesser General Public Licence version 3 (LGPL Version 3).
\end{quote}
\item For software distributed under the Apache licence v2\cite{[11]}:
\begin{quote}
This software is distributed under the terms of the Apache version 2.0 licence.
\end{quote}
\item For software distributed under the BSD-2-Clause licence\cite{[12]}:
\begin{quote}
This software is distributed under the terms of the BSD-2-Clause licence.
\end{quote}
\item For software distributed under the BSD-3-Clause licence\cite{[13]}:
\begin{quote}
This software is distributed under the terms of the BSD-3-Clause licence.
\end{quote}
\item For software distributed under the MIT licence\cite{[14]}:
\begin{quote}
This software is distributed under the terms of the MIT licence.
\end{quote}
\end{itemize}
